---
title: "P8105_HW1_msp2210"
output: github_document
author: "Mukta Patwari"
date: "2025-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

**Loading libraries**

```{r}
library(tidyverse)
library(moderndive)
```

**Loading the dataset**

```{r}
data("early_january_weather")
```

This dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. This dataset has 15 variables, which are `r names(early_january_weather)`. The mean temperature is `r mean(pull(early_january_weather, temp))` degrees Fahrenheit. The highest temperature recorded was `r max(pull(early_january_weather, temp))` degrees Fahrenheit and the lowest temperature recorded was `r min(pull(early_january_weather, temp))` degrees Fahrenheit.

**Creating a scatterplot**

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color=humid)) + geom_point()
```

Based on this plot, it is clear how temperature fluctuates throughout the day. There is a  zig-zag effect, because higher temperatures are occurring during the day and then they decline at night (and this is a cycle that continues throughout the month). The beginning of the month has cooler temperatures during daytime and nighttime, and by January 14th, both of these temperatures have increased. Also, there was less humidity at the beginning of the month when compared to the days leading up to January 14th.

**Saving scatterplot to directory**

```{r}
ggsave("p8105_hw1_msp2210_scatterplot.pdf", height = 4, width = 6)
```
